---
title: "general notes"
author: Rakesh Poduval, `r Sys.Date()`
output:
  html_document: 
    toc: TRUE
    toc_depth: 2
    toc_float: TRUE
    theme: yeti
    highlight: monochrome
---

Documenting few notes on R which I think is useful.

# Changing language
Say the default language is something else and you want to make it english. Then Copy the following file to desktop : `C:\Program Files\R\R-3.0.2\etc\Rconsole` and then at the very bottom of it put language = en. Save, close and replace it in the above destination with the previous one, you are done.

# Download R packages
To download all R packages available and put it into your personal folder you can execute the following scripts. Note that it downloads all packages available in cran, thus takes a lot of time. 
```{r }
if (F) { 
  pkg.list = available.packages()
  download.packages(pkgs = pkg.list, 
                    destdir = "put destination folder path here")
}
```

# Packages installed 
To view the R packages installed in your R session run the following scripts.
```{r }
if (F) { 
  ip <- installed.packages()
  names(ip[,'Package'])
}
```

# Extract inbuild functions
The internal functions could be hacked sometimes like this
```{r }
if (F) {
  getAnywhere(boxplot.default) 
  mthods()
}
```

# Get system information
Basic system information cen be obtained like this.
```{r }
Sys.info() # system information
sessionInfo() # about R and the session information
```

# Colors in R
The colors avaible in R are 
```{r }
RColorBrewer:::display.brewer.all()
```
The following script can be useful to visualize the colors which are chosen.
```{r }
col <- c("#D78113", "#E0A060", "#E0C0A0", "#B0C0D0", 
         "#80B0C0", "#6090B0", "#307090", "#005684");
plot(0, 
     0, 
     type = "n", 
     xlim = c(0, 1), 
     ylim = c(0, 1),
     axes = F, 
     xlab = "", 
     ylab = "")
rect(0:(length(col) - 1)/length(col), 
     0, 
     1:length(col)/length(col), 
     1, 
     col = col)
```

# R fonts
The following scripts are useful to view the avaibale windows fonts. You can also get additional fonts into R.
```{r }
# list of available fonts
windowsFonts() 

# adding additional fonts
windowsFonts(
  arial   = windowsFont("Arial Black"),
  bookman = windowsFont("Bookman Old Style"),
  comic   = windowsFont("Comic Sans MS"),
  calibri = windowsFont("Calibri"),
  courier = windowsFont("Courier New")
)

# Checking the fonts
plot(1:10, 1:10, type = "n")
for (i in 1:length(windowsFonts())) {
  text(5, i + 1, family = names(windowsFonts())[[i]], windowsFonts()[[i]])
}

```
