---
title: "R Notebook"
output: html_notebook
---

# Package DT features

Let's use `iris` data set for demonstration.

Firs we need initialize the R environment.

```{r init, echo=FALSE, message=FALSE}
library(DT)
library(dplyr)
```

Normal print (`dplyr`) method (this is pretty well as well)

```{r}
iris
```

## About

The R package DT provides an R interface to the `JavaScript` library `DataTables`. R data objects (matrices or data frames) can be displayed as tables on HTML pages, and `DataTables` provides filtering, pagination, sorting, and many other features in the tables.

The default options comes with various features such as:

-   Number of entries to be displayed in each page

-   Search option (performs text search)

-   Sorting columns etc.

Let us demonstrate the `DT:::datatable()` features.

```{r}
iris %>% datatable()
```

`datatable()` has the following arguments. Please execute `help(datatable)` for more information.

```{r}
args(datatable)
```

**We try to explain each of these parameters with simple examples**

## Controlling parameters

It is possible to disable these default viewing options.

```{r}
iris %>% head %>% 
  datatable(options = list(searching = FALSE, 
                           lengthChange = FALSE, 
                           lengthMenu = c(2, 4, 6, 8),
                           ordering = FALSE,
                           info = FALSE, 
                           paging = FALSE,
                           pageLength = 4))
```

**Remark :** With `editable = TRUE`, you can enable editing.

## Table formatting features

Before proceeding we fix some options

```{r}
display_options <- list(searching = FALSE, 
                        lengthChange = FALSE, 
                        ordering = FALSE,
                        info = FALSE, 
                        paging = FALSE,
                        pageLength = 4)
```